Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-28 paper/paper.tex
resolving conflicts

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks.}
  \centering
  \begin{tabular}{llllp{10cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Type} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule     
    Ping Scan & NMAP & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & hping3 & DoS & TCP & \textbf{flags}: S,12;
    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5;\\
    \midrule
    JBoss JMX\MyComment{ Console Beanshell Exploit} & Nikto2 & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    WeBid \MyComment{Local File Inclusion} & RFI & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    ColdFusion \MyComment{PESC} & Nikto2 & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    .htaccess access\MyComment{attempted-recon} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    iisadmin access\MyComment{web-application-attack} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & Nikto2 & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & Nikto2 & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    Apache Struts 2 & \Fix{Source} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\
    WordPress Login & \Fix{Source} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

    \midrule
    Ping Scan & 5 & 5 & 2 & - \\
    SYN Flood & 3 & 2 & 2 & - \\
    JBoss JMX & 4 & 23 & 228 & \Fix{23} \\
    WeBid & 1 & 23 & - & -\\    
    ColdFusion & 6 & 19 & 34 & \Fix{5}\\
    .htaccess access & 4 & 17 & 984 & \Fix{2}\\
    .idq access & 13 & 19 & 511 & \Fix{2} \\
    iisadmin access & 3 & 17 & \Fix{861} & \Fix{1} \\
    /system32/ in URI & 3 & 20 & 1,913 & \Fix{2} \\
    Script tag in URI & 22 & 22 & 786 & - \\
    Apache Struts 2 & - & - & - & - \\
    Wordpress Login & - & - & - & - \\
    \bottomrule
  \end{tabular}
\end{table}

@@

%% It is important to note that this setup is \emph{not}
%% necessary if \emph{pcap} files for the attack already exist. We
%% proceeded as follows to obtain pcap files for the the attacks we
%% analyzed: we 1)~instantiated the script (\eg{}, defined the IP of the
%% web server), 2)~executed the script, 3)~and monitored the traffic with
%% the Wireshark network monitor. The
%% Wireshark tool records the traffic and reports pcap files on output
%% for replaying it.

\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
pcap files on input. As explained on Section~\ref{sec:minimization},
these files abstract network messages~\cite{pcap} and can be used to
avoid network traffic during rule synthesis. For some of the attacks
listed on Table~\ref{table:attacks}, we obtained pcap files directly
from trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}\Fix{confirm}).\Mar{reminder: we still
  don't have any example of attack listed on Table~\ref{table:attacks}
  whose pcap file was obtained from those websites.} \Luc{Ataques Apache 
  e WordPress. @Guilherme adicionar fontes} In addition to
these sources, we used the Nikto2~\cite{nikto} open source security
tool as follows to obtain pcap files for (non-flooding) http and tcp
attacks. We configured an http server (Apache2) and ran Nikto2 to
attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks. After
execution, we inspected the \suri\ log to identify which rules have
been activated and which packets were involved on each
activation. With the packet numbers, it is possible to recover the
corresponding pcap files, using Wireshark, as to ran \tname{} on
them. It is worth noting that some rules are triggered multiple times,
indicating that Nikto2 expresses multiple variants of the same
attack. RQ3 evaluates the importance of these variants to \tname{}'s
performance. We used similar methodology to obtain pcap files using
other security testing tools. We used NMAP~\cite{netmap} to reproduce
the Ping Scan attack (see Section~\ref{sec:active-recon}) and
hping3~\cite{hping3} to reproduce the SYN Flood attack (see
Section~\ref{sec:dos}).

%% \Mar{@Lucas, diga exatamente como vc. obtem as variacoes. isto eh que
%%   eh importante para a proxima sentenca.}.  \Luc{Eh feito da mesma
%%   forma q foi descrita acima. Quando um ataque produz varios pacotes
%%   que geram alertas, nos criamos um pcap com um dos pacotes e
%%   utilizamos ele como para gerar a regra maxima inicial.  Os outros
%%   pacotes sao salvos em um outro pcap e sao utilizados para calcular o
%%   recall.}

%% \Mar{what about the variants? where do
%%   they come from? this is confusing as the narrative above suggests
%%   the attack is hardcoded in the tool.}\Luc{O trafego de um ataque
%%   normalmente possui varios pacotes, entre esses varios pacotes alguns
%%   deles possuem um padrao. As regras sao criadas descrevendo esse
%%   padrao, com isso um unico ataque pode disparar varios alertas.

%%   Porem, existem ataques q so produzem um alerta, ou seja, a regra criada 
%%   so detecta um dos pacotes do trafego. Nesses casos, procuramos na internet 
%%   datasets (arquivos pcap) que contivessem mais pacotes desses ataques.
%%   Fizemos isso tambem para ataques q nos ja tinhamos variacoes, pois
%%   (teoricamente) quanto mais variacoes, melhor. Existem alguns sites onde 
%%   eh possivel encontrar pcaps com o trafego de alguns ataques, n consegui
%%   adicionar nas referencias: 
%%   https://www.netresec.com/?page=PcapFiles
%% }

%% According to the TcpReplay web
%% site ``This capture is much larger and has a smaller average packet
%% size than the previous capture (smallFlows.pcap). It also has many
%% more flows and different applications.''

\subsubsection{Benign traffic.}~Public data sets of bening traffic
exist; they are important to evaluate the rate of false positives of
security tools. We used two popular public sources. The Bigflows.pcap
data set is part of the TcpReplay~\cite{tcpreplay} open source
project. It includes real network traffic of a busy private network
access point to the Internet. In addition to the Bigflows.pcap data
set, we also used data sets\Mar{more than one?}  made available by
Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
  of the traffic? as the description for
  Bigflows.pcap?}. \Luc{@Guilherme ...}

>>>>

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks. Non-http attacks
    appear at the top andcontent-based http attacks appear at the bottom.}
  \centering
  \begin{tabular}{llllp{10cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Type} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule
    Black Nurse & \cite{pcap-attacks} & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \textbf{threshold}: type both, track by\_dst, count 250, seconds 1;\\    
    Ping Scan & \cite{netmap} & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & \cite{hping3} & DoS & TCP & \textbf{flags}: S,12;
    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5;\\
    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
    UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5; \\
    \midrule
    Apache Struts 2 & \cite{wrccdc} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\    
    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    WeBid \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

    \midrule
    Black Nurse & - & - & - & - \\    
    Ping Scan & 6 & 5 & 2 & - \\
    SYN Flood & 4 & 2 & 2 & - \\
    UDP Teardrop & - & - & - & - \\
    UDP Flood & - & - & - & - \\
    \midrule
    Apache Struts 2 & - & - & - & - \\
    ColdFusion & 7 & 19 & 34 & \Fix{5}\\
    iisadmin access & 4 & 17 & \Fix{861} & \Fix{1} \\        
    JBoss JMX & 5 & 23 & 228 & \Fix{23} \\
    Script tag in URI & 23 & 22 & 786 & - \\
    WeBid & 1 & 23 & - & -\\    
    Wordpress Login & 8 & 59 & - & - \\
    .htaccess access & 5 & 17 & 984 & \Fix{2}\\
    .idq access & 14 & 19 & 511 & \Fix{2} \\
    /system32/ in URI & 4 & 20 & 1,913 & \Fix{2} \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\vspace{1ex}
\subsubsection{\label{subsec:malicious-traffic}Malicious traffic.}~Recall that \tname\ takes
pcap files on input. As explained on Section~\ref{sec:minimization},
these files abstract network messages~\cite{pcap} and can be used to
avoid network traffic during rule synthesis. For some of the attacks
listed on Table~\ref{table:attacks}, we obtained pcap files directly
from trusted security websites (\eg{},
NetReseC~\cite{pcap-attacks}). In addition to these sources, we used
the Nikto2~\cite{nikto} open source security tool as follows to obtain
pcap files for some of the http attacks. We configured an http server
(Apache2) and ran Nikto2 to attack that server. Simultenously, we ran
WireShark~\cite{wireshark-net-monitor} to record the network traffic
and Suricata to analyze the traffic. We configured \suri\ with all
rules from the ``Emerging Threats Open
Ruleset''~\cite{emerging-threats-open} and configured Nikto2 with
default parameters, which makes it replay all recorded attacks from
its benchmark. After execution, we inspected the \suri\ log to
identify which rules have been activated and which packets were
involved on each activation. With the packet numbers, it is possible
to recover the corresponding pcap files, using Wireshark, as to ran
\tname{} on them separately. It is worth noting that some rules are
triggered multiple times, indicating that Nikto2 expresses multiple
variants of the same attack. RQ3 evaluates the importance of these
variants to \tname{}'s performance. We used similar methodology to
obtain pcap files using other security testing tools. For example, we
used NMAP~\cite{netmap} to reproduce the Ping Scan attack (see
Section~\ref{sec:active-recon}) and hping3~\cite{hping3} to reproduce
the SYN Flood attack (see Section~\ref{sec:dos}).

\vspace{1ex}
\subsubsection{Benign traffic.}~Public data sets of bening traffic
exist; they are important to evaluate the rate of false positives of
security tools. We used two popular public sources. The Bigflows.pcap
data set is part of the TcpReplay~\cite{tcpreplay} open source
project. It includes real network traffic of a busy private network
access point to the Internet. We also used the

\Fix{---------------- estou aqui}

In addition to the Bigflows.pcap data
set, we also used data sets\Mar{more than one?}  made available by
Stratosphere Lab~\cite{stratosphere-normal}\Mar{is there a description
  of the traffic? as the description for
  Bigflows.pcap?}. \Luc{@Guilherme ...} \Gui{The Stratosphere IPS Project has a sister project called the Malware Capture Facility Project that is responsible for making the long-term captures. This project is continually obtaining malware and normal data to feed the Stratosphere IPS (copy paste do site). No site tem uma lista de datasets, e embaixo dele descricao da origem desses datasets, como de uma rede universitaria, um usuario normal, etc. https://www.stratosphereips.org/datasets-normal}
